Mitigating Interest Flooding DDoS Attacks in NDN

### 2013-01-12 04a-design-naive.tex
Update design section
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@900 df1a90cd-247b-4364-bdaf-287a704eb499

An obvious and na\"ive solution to defend against Interest flooding attacks is to restrict the number of Interests that are forwarded in the network. To this end, we exploit a 
 fundamental principle of NDN architecture---flow balance between Interest and Data packets. The flow balance refers to the fact that one Interest can be satisfied by at most one Data packet. This principle allows intermediate routers that forward the Interest packets to control the number of outstanding Interests in the network. 
One simple implementation technique  is for an NDN router to limit the number of forwarded Interests out of each interface based on the physical capacity of the corresponding out interface. This technique is a slight modification of the well-known {\it Token Bucket} algorithm that is currently widely used in packet switched networks. Analogous to the {\it Token Bucket} algorithm, NDN routers keep track of the amount of Data requested that can fully utilize the downstream link (estimated from the number of Interests it forwarded) and once the link capacity limit has been reached, they no longer forward new incoming Interests. Ideally, the number of tokens {\it i.e.} the pending \emph{Interests Limit} for each link will be proportional to the link's bandwidth-delay product (BDP)~\cite{tcp-survey}. We can formalize this value as follows:
% With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:
%
\begin{equation}
\small \mathrm{Interest\ Limit} = Delay\ [s] \cdot \frac{\mathrm{Bandwidth\ [Bytes/s]}}{\mathrm{Data\ packet\ size\ [Bytes]}}
\end{equation}

In the above equation, \emph{Delay} is the expected time of Interest being satisfied and \emph{Data packet size} is the size of the returned Data packet.
Although both these values are not known {\it apriori}, it is not really necessary to use their exact values.
One can simply set the pending Interest limit based on the average values of round trip time and observed Data packet size, as the network buffers can smooth out most of the network fluctuations.

This {\it Token Bucket} approach might be exceptionally retrictive in forwarding Interests -- not all Interests will result in a Data packet -- thus underutilizing the network. However, the biggest drawback of the algorithm is that it can be used as an effective mechanism to carry out denial of service attacks. If a router has utilized all its {\it tokens} to forward {\it bad} Interests on an out Interface, it can no longer forward incoming Interests from legitimate users on that out inteface till the pending {\it bad} Interests start to expire. One way to get around this issue is to impose a per-interface limit so that {\it bad} Interests are not allowed to entirely consume the limits of a specific interface. We describe this technique in greater detail below.

@@

The key drawback of the {\it Token bucket with per interface fairness} algorithm is that it still admits Interests from malicious users. A sizeable percentage of these malicious Interests are forwarded all the way to content producers, thereby utilizing scarce network resources and reducing resources available to serve legitimate users.  This algorithm attempts to ensure that each interface does not forward more than its fair share of Interests, but in doing so, it drops both legitimate Interests as well as malicious Interests. For any strategy to be effective in defending against Interest flooding attacks, it must be able to detect and effectively drop only malicious requests for content. `Good' Interests from legitimate users must be admited and forwarded appropriately. The key question is how can we devise mitigation algorithms that allow a router to distinguish between `good' and `bad' Interests?

Unlike in today's host-based Internet architecture where DDoS mitigation techniques that rely on blacklisting of dark IP prefixes or whitelisting of good IP prefixes are still reasonably effective, in NDN one cannot easily distinguish a `good' prefix from a `bad' one. Content producers must be able to support dynamic generation of Data packets to satisfy incoming Interest packets -- thus, an Interest that cannot be satisfied by matching Data from any intermediate router's cache is necessarily forwarded to the producer for possible dynamic Data packet generation. Thus intermediate routers cannot determine if the prefix is `good' or `bad' based on the name alone -- DDoS mitigation techniques that rely on whitelisting or blacklisting of certain namespaces  will be inefficient and likely ineffective.

>>>>

An obvious and na\"ive solution to defend against Interest flooding attacks is to restrict the number of Interests that are forwarded in the network. To this end, we exploit a 
 fundamental principle of NDN architecture---flow balance between Interest and Data packets. The flow balance refers to the fact that one Interest can be satisfied by at most one Data packet. This principle allows intermediate routers to control the inbound Data traffic by controlling the number of outstanding Interests in the network. 
One simple implementation technique  is for an NDN router to limit the number of forwarded Interests out of each interface based on the physical capacity of the corresponding out interface. This technique is a slight modification of the well-known {\it Token Bucket} algorithm that is currently widely used in packet switched networks. Analogous to the {\it Token Bucket} algorithm, NDN routers keep track of the amount of Data requested that can fully utilize the downstream link (estimated from the number of Interests it forwarded) and once the link capacity limit has been reached, they no longer forward new incoming Interests. Ideally, the number of tokens (the pending \emph{Interests Limit}) for each link will be proportional to the link's bandwidth-delay product (BDP)~\cite{tcp-survey}. We can formalize this value as follows:
% With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:
%
\begin{equation}
\small \mathrm{Interest\ Limit} = Delay\ [s] \cdot \frac{\mathrm{Bandwidth\ [Bytes/s]}}{\mathrm{Data\ packet\ size\ [Bytes]}}
\end{equation}

In the above equation, \emph{Delay} is the expected time of Interest being satisfied and \emph{Data packet size} is the size of the returned Data packet.
Although both these values are not known a priori, it is not really necessary to use their exact values.
One can simply set the pending Interest limit based on the average values of round trip time and observed Data packet size, as the network buffers can smooth out most of the network fluctuations.

This {\it Token Bucket} approach might be exceptionally retrictive in forwarding Interests---not all Interests will result in a Data packet---thus underutilizing the network. However, the biggest drawback of the algorithm is that it can be used as an effective mechanism to carry out denial of service attacks. If a router has utilized all its tokens to forward malicious Interests, it can no longer forward incoming Interests from legitimate users till the pending malicious Interests start to expire. One way to get around this issue is to impose a per interface limit, so that malicious Interests are not allowed to entirely consume the limits of a specific interface. We describe this technique in greater detail below.

@@

The key drawback of the {\it Token bucket with per interface fairness} algorithm is that it admits Interests from malicious users. A sizeable percentage of these malicious Interests are forwarded all the way to content producers, thereby utilizing scarce network resources and reducing resources available to serve legitimate users.  This algorithm attempts to ensure that each interface does not forward more than its fair share of Interests, but in doing so, it drops both legitimate Interests as well as malicious Interests. For any strategy to be effective in defending against Interest flooding attacks, it must be able to detect and effectively drop only malicious requests for content. `Good' Interests from legitimate users must be admited and forwarded appropriately. The key question is how can we devise mitigation algorithms that allow a router to distinguish between `good' and `bad' Interests? 
One can try a black- or whitelisting approach. However, besides the fact that source black- and whitelisting cannot work in NDN as it does not feature source addresses, it requires extraneous knowledge to classify the incoming Interests.

% Alex: I'm not sure that this part is relevant... The point I see is that blacklisting and whitelisting is not 100% effective, but it can be used in both worlds

% In today's host-based Internet architecture where DDoS mitigation techniques rely on blacklisting of ``dark'' IP prefixes or whitelisting of good IP prefixes.
% Unlike IP, in NDN one cannot easily distinguish a `good' prefix from a `bad' one. 
% Content producers must be able to support dynamic generation of Data packets to satisfy incoming Interest packets---thus, an Interest that cannot be satisfied by matching Data from any intermediate router's cache is necessarily forwarded to the producer for possible dynamic Data packet generation. 
% Thus intermediate routers cannot determine if the prefix is `good' or `bad' based on the name alone---DDoS mitigation techniques that rely on whitelisting or blacklisting of certain namespaces  will be inefficient and likely ineffective.

### 2013-01-12 algos/dynamic-limits.tex
Update design section
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@900 df1a90cd-247b-4364-bdaf-287a704eb499

The previous strategy---the satisfaction-based Interest acceptance---divides the available forwarding tokens among all interfaces in proportion to their Interest satisfaction ratios.
% Given the drawbacks of this algorithm, the same effect can be achieved without relying on probabilistic models. 
An alternate strategy, which does not have the described drawbacks, is to enable and enforce explicit Interest limit for each incoming interface, where the value of the limit depends directly on the interface's Interest satisfaction ratio.
Routers need to announce these limits to their downstream neighbors, ensuring that any Interest forwarded from the downstream router is allowed to get through, resulting in genuine Interest satisfaction statistics.

The formal definition of the satisfaction-based pushback algorithm is presented in Pseudocode~\ref{alg:dynamic limits}, while Fig.~\ref{fig:dynamic limits example} illustrates how the algorithm would work in our example in Fig.~\ref{fig:flooding example}.
Assuming an initial token bucket limit $L=10$ and the current satisfaction ratio for router A is 50\% for \texttt{eth1} and 0\% for \texttt{eth0}, and for router B the ratio is 30\% for \texttt{eth0}, each node would set and announce the following  incoming interface limit $L'$: 
\begin{enumerate}
\item router B would set and announce the incoming interface limit $L'=3$;
\item router A, after receiving announcement from B would readjust its incoming interface limits to $L'_{eth1} = 1.5$ and $L'_{eth0} = 0$; and
\item both legitimate users and adversaries may either obey or ignore the announced limit, which will be in any case enforced by router A.
\end{enumerate}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{dynamic-limits}
  \caption{Satisfaction-based pushback example: routers explicitly tell neighbors how many Interest packets they can deliver to the Data producer}
  \label{fig:dynamic limits example}
\end{figure}

@@

We note that while in the description of the satisfaction-based pushback algorithm we explicitly used ``outgoing'' and ``incoming'' interfaces,  all interfaces can be both incoming and outgoing.
Thus, it may not be entirely clear which outgoing limit $L_{out}$ (line 10 in the algorithm) should be used to calculate the incoming limit $L_{in}$.
To overcome this problem, in our actual implementation we enforced separate incoming/outgoing interface limits for each individual FIB entry.
That is, for each FIB entry we set a separate Interest limit for each incoming interface ($L_{in}^{'FIB}$) based on a sum of FIB entry limits for each outgoing interface $L=\sum{L_{out}^{FIB}}$.

% Alex: not sure if this paragraph belongs here
Essentially, both satisfaction-based Interest acceptance and satisfaction-based pushback algorithms are forms of a well-known push-back mechanism~\cite{Pushback}, with several core differences. First, we are suppressing (pushing back) unwanted requests for Data, not actual Data itself.
Second, differentiating between good and bad Interests is based on the traffic symmetry principle of NDN.
% Alex: I'm not entirely sure about this point... 
Finally, both intelligent attack mitigation algorithms can be deployed at all times without degrading network performance even when there are no active attackers.

>>>>

The previous algorithm---the satisfaction-based Interest acceptance---divides the available forwarding tokens among all interfaces in proportion to their Interest satisfaction ratios.
% Given the drawbacks of this algorithm, the same effect can be achieved without relying on probabilistic models. 
An alternate algorithm for proportional token distribution without overreaction is to enable and enforce explicit Interest limit for each incoming interface, where the value of the limit depends directly on the interface's Interest satisfaction ratio.
Routers need to announce these limits to their downstream neighbors, ensuring that any Interest forwarded from the downstream router is allowed to get through, resulting in genuine Interest satisfaction statistics.

The formal definition of the satisfaction-based pushback algorithm is presented in Pseudocode~\ref{alg:dynamic limits}, while Fig.~\ref{fig:dynamic limits example} illustrates how the algorithm will work in our example in Fig.~\ref{fig:flooding example}.
Assuming an initial token bucket limit $L=10$ and the current satisfaction ratio for router A is 50\% for \texttt{eth1} and 0\% for \texttt{eth0}, and for router B the ratio is 30\% for \texttt{eth0}, each node will set and announce the following  incoming interface limit $L'$: 
\begin{enumerate}
\item router B will set and announce the incoming interface limit $L'=3$;
\item router A, after receiving announcement from B will readjust its incoming interface limits to $L'_{eth1} = 1.5$ and $L'_{eth0} = 0$; and
\item both legitimate users and adversaries may either obey or ignore the announced limit, which will be in any case enforced by router A.
\end{enumerate}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{dynamic-limits}
  \caption{Satisfaction-based pushback example
%: routers explicitly tell neighbors how many Interest packets they can deliver to the Data producer%
}
  \label{fig:dynamic limits example}
\end{figure}

@@

We note that while in the description of the satisfaction-based pushback algorithm we explicitly used ``outgoing'' and ``incoming'' interfaces,  all interfaces can be both incoming and outgoing.
Thus, it may not be entirely clear which outgoing limit $L_{out}$ (line 10 in the algorithm) should be used to calculate the incoming limit $L_{in}$.
To overcome this problem, in our actual implementation we enforced separate incoming/outgoing interface limits for each individual FIB entry.
That is, for each FIB entry we set a separate Interest limit for each incoming interface (${L'}_{in}^{fib}$) based on a sum of FIB entry limits for each outgoing interface $L=\sum{L_{out}^{fib}}$.

Both satisfaction-based Interest acceptance and satisfaction-based pushback algorithms are forms of a well-known push-back mechanism~\cite{Pushback}, but with several core differences. 
First, we are suppressing (pushing back) unwanted requests for Data, not actual Data itself.
Second, differentiating between good and bad Interests is based on the traffic symmetry principle of NDN.
% Alex: I'm not entirely sure about this point... 
Finally, both intelligent attack mitigation algorithms can be deployed at all times without degrading network performance even when there are no active attackers.

### 2013-01-12 algos/fair-queueing.tex
Small changes
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@899 df1a90cd-247b-4364-bdaf-287a704eb499
Update design section
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@900 df1a90cd-247b-4364-bdaf-287a704eb499

To address the lack of fairness associated with the na\"ive {\it Token Bucket} algorithm, we modify it to ensure that the Interests forwarded by a router on each interface represent a fair mix of Interests received from neighboring interfaces. For example, in Fig.~\ref{fig:flooding example} router A can ensure that the tokens associated with Interests sent out on interface {\texttt eth2}  are fairly distributed across incoming interfaces \texttt{eth0} and \texttt{eth1}. 
%Due to the expected small volume of Interests, we cannot merely rely on network buffers to perform statistical multiplexing of Interests, as they would almost never be %buffered.
%At the same time, until bag of tokens is not empty, there is no reason to delay Interest forwarding, as we do not known how many and from which interfaces Interests %will arrive in the future.
In order to achieve our goal of ensuring ``fair'' mixing of Interests from all neighboring interfaces, 
%, we implement an additional features to support buffering of Interests, if they cannot be immediately forwarded. 
we extend the Pending Interest Table to support flagging of Interests that cannot be immediately forwarded (see Fig.~\ref{fig:queueing}). To support fair queuing, we implement hierarchical queues for each interface as shown in Fig.~\ref{fig:queueing})\footnote{This mechanism is essentially class based queuing, with classes for each outgoing/incoming interface.} 
%As an alternative to hierarchical queues, one can also use an approach based on virtual time. For details, please see ~\cite{zhang1990virtual}.
We note that unlike normal queuing, Interest queues do not actually store a packet, but merely a bi-directional pointer to the existing PIT entry.
Thus, a PIT entry can be quickly updated when the Interest is actually forwarded; further the element can be quickly removed from the queue when the Interest expires.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{queue}
  \caption{Interest queuing: if tokens are unavailable, the router creates a PIT entry, but instead of forwarding, it enqueues the Interest}
  \label{fig:queueing}
\end{figure}

We present a formalized description of this algorithm in Pseudocode~\ref{alg:queuing}. We can control the amount of physical resources utilized at a router (such as memory and computation power) by setting appropriate queue sizes. It is also important to set apprpriate value for how long an Interest can be enqueued. If an Interest is enqueued for a long time, by the time it is dequeued and appropriately forwarded, the resulting Data packet might not be valid due to state expiration at previous downstream routers. We believe that implementing additional mechanisms on each pair of communicating routers to keep the Interests from getting stale would mitigate this issue, though the details are beyond the scope of this paper.
{\color{red} \it Priya:  We need to  touch upon the Interest freshness issue in the NDN overview section; else we should get rid of this bit.}

@@

As we show in Section~\ref{sec:evaluation}, this algorithm provides partial relief from Interest flooding attacks, allowing legitimate users to successfully fetch Data for 15--20\% of their expressed Interests. While this algorithm might be reasonable for ensuring fairness in an NDN network, it is ineffective in protecting legitimate users from malicious attackers. Malicious attackers are able to successfully thwart access to content for legitimate users by sending a relatively modest volume of Interests.

>>>>

To address the lack of fairness associated with the na\"ive {\it Token Bucket} algorithm, we modify it to ensure that the Interests forwarded by a router on each interface represent a fair mix of Interests received from neighboring interfaces. For example, in Fig.~\ref{fig:flooding example} router A can ensure that the tokens associated with Interests sent out on interface {\texttt eth2}  are fairly distributed across incoming interfaces \texttt{eth0} and \texttt{eth1}. 
%Due to the expected small volume of Interests, we cannot merely rely on network buffers to perform statistical multiplexing of Interests, as they would almost never be %buffered.
%At the same time, until bag of tokens is not empty, there is no reason to delay Interest forwarding, as we do not known how many and from which interfaces Interests %will arrive in the future.
In order to achieve our goal of ensuring ``fair'' mixing of Interests from all neighboring interfaces, 
%, we implement an additional features to support buffering of Interests, if they cannot be immediately forwarded. 
we extend the Pending Interest Table to support flagging of Interests that cannot be immediately forwarded and implement hierarchical queues for each interface (see Fig.~\ref{fig:queueing}). 
This mechanism is essentially a class based queuing~\cite{floyd1995link}, with classes for each outgoing/incoming interface.
%As an alternative to hierarchical queues, one can also use an approach based on virtual time. For details, please see ~\cite{zhang1990virtual}.
We note that unlike normal queuing, Interest queues do not actually store a packet, but merely a bi-directional pointer to the existing PIT entry.
Thus, a PIT entry can be quickly updated when the Interest is actually forwarded; further the element can be quickly removed from the queue when the Interest expires.

@@

\begin{figure}[thb]
  \centering
  \includegraphics[scale=0.65]{queue}
  \caption{Interest queuing: if tokens are unavailable, the router creates a PIT entry, but instead of forwarding, it enqueues the Interest}
  \label{fig:queueing}
\end{figure}

We present a formalized description of this algorithm in Pseudocode~\ref{alg:queuing}. 
By setting appropriate queue sizes we can control the amount of physical resources utilized at a router (such as memory and computation power). 
It is also important to set apprpriate value for how long an Interest can be enqueued. 
If an Interest is enqueued for a long time, by the time it is dequeued and appropriately forwarded, the resulting Data packet might not be valid due to state expiration at the downstream routers. 
For our evaluations we chose to enqueue Interests up to 10\% of their original lifetime.
% We believe that implementing additional mechanisms on each pair of communicating routers to keep the Interests from getting stale would mitigate this issue, though the details are beyond the scope of this paper.
% Alex: I don't believe there is a need for separate discussion about "freshness", as we already introduced concept of lifetime
% {\color{red} \it Priya:  We need to  touch upon the Interest freshness issue in the NDN overview section; else we should get rid of this bit.}

@@

As we show in Section~\ref{sec:evaluation}, this algorithm provides partial relief from the Interest flooding attacks, allowing legitimate users to successfully fetch Data for 15--20\% of their expressed Interests. 
We note that while this algorithm might be reasonable for ensuring fairness in an NDN network, it is ineffective in protecting legitimate users from malicious attackers. 
Malicious attackers are able to successfully thwart access to content for legitimate users by sending a relatively modest volume of Interests.

### 2013-01-12 algos/probabilistic.tex
Update design section
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@900 df1a90cd-247b-4364-bdaf-287a704eb499

Parameter $\theta$ on line 5 of the Pseudocode~\ref{alg:probabilistic model} ensures that the probabilistic model is not enforced when the volume of Interests arriving at a particular interface is small. This step is critical---while we want to drop Interests from attackers, we also want to provide an opportunity for legitimate users to regain their share of resources after temporary Data delivery failures due to  congestion.

>>>>

Parameter $\theta$ on line 5 of the Pseudocode~\ref{alg:probabilistic model} ensures that the probabilistic model is not enforced when the volume of Interests arriving at a particular interface is small. This step is critical---while we want to drop Interests from attackers, we also want to provide an opportunity for legitimate users to regain their share of resources after temporary Data delivery failures.
